---
title: "Your Topic Title"
author: "Phani Srija Kesara"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    toc: yes
    toc_float: yes
---

```{r include = FALSE}

# SET GLOBAL KNITR OPTIONS

knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 8)

# LOAD PACKAGES

library(pander)
library(kableExtra)
library(forecast)
library(zoo)

```

<br>
<br>

# Introduction

This code through explores the `AirPassengers` dataset, which records the monthly totals of international airline passengers (in thousands) from 1949 to 1960. This dataset provides a great example for demonstrating time series analysis techniques, including trend analysis, seasonality, and decomposition.

<br>

## Content Overview

Specifically, we'll explain and demonstrate how to manipulate time series data, visualize the dataset, and perform exploratory analysis using R. We will further analyze trends and seasonality, which are crucial for understanding long-term behavior in datasets such as this one.

<br>

## Why You Should Care

This topic is valuable because understanding time series analysis is essential in many fields, such as finance, economics, and business. Analyzing historical data helps us forecast future trends, detect anomalies, and make informed decisions.

<br>

## Learning Objectives

Specifically, you'll learn how to:
1. Load and inspect time series data in R.
2. Visualize time series data using `ggplot2`.
3. Decompose a time series into trend, seasonality, and residuals.
4. Use basic forecasting methods.

<br>
<br>

# Body Title

Here, we'll show the initial steps to load and visualize the dataset. This includes loading the data, converting it into a format suitable for analysis, and performing basic exploratory data analysis.

<br>

## Further Exposition

This analysis is based on the foundational concepts of time series analysis, as developed and used in many practical applications. The dataset and methods used will be further explored to provide a deeper understanding of seasonal patterns and trends in air travel. 

<br>

## Basic Example

A basic example shows how to load, visualize, and explore the AirPassengers dataset to observe trends over time.

```{r}

# Load the AirPassengers dataset
data("AirPassengers")

# View the first few values of the dataset
head(AirPassengers)

# A basic plot of the data
plot(AirPassengers, 
     main = "AirPassengers Data: 1949-1960", 
     ylab = "Number of Passengers", 
     xlab = "Time", 
     col = "darkgreen", 
     lwd = 2, 
     type = "o")  # 'type = "o"' shows both points and lines

```

<br>

## Advanced Examples

More specifically, this can be used for calculating and visualizing a moving average to smooth out short-term fluctuations and highlight long-term trends in the AirPassengers dataset.

```{r}
# Load the AirPassengers dataset
data("AirPassengers")

# Calculate the 12-month moving average
moving_avg <- stats::filter(AirPassengers, rep(1/12, 12), sides = 2)

# Plot the original data and the moving average
plot(AirPassengers, 
     main = "AirPassengers with 12-Month Moving Average", 
     ylab = "Number of Passengers", 
     xlab = "Time", 
     col = "blue", 
     lwd = 2)

# Add the moving average to the plot
lines(moving_avg, col = "red", lwd = 2)

# Add a legend to distinguish between the original data and moving average
legend("topleft", 
       legend = c("Original Data", "12-Month Moving Average"), 
       col = c("blue", "red"), 
       lty = 1, 
       lwd = 2)
```

<br>

What's more, it can also be used for forecasting future values using models like HoltWinters to predict trends and seasonal patterns for the next 12 months based on historical data.

```{r}

# Fit the HoltWinters model
hw_model <- HoltWinters(AirPassengers)

# Forecast the next 12 months (1 year)
forecasted_values <- forecast::forecast(hw_model, h = 12)

# Plot the original data and forecasted values
plot(forecasted_values, 
     main = "AirPassengers Forecast Using HoltWinters Model", 
     xlab = "Year", 
     ylab = "Number of Passengers")

# Add original data line for comparison
lines(AirPassengers, col = "blue", lwd = 2)

```

<br>

Most notably, it's valuable for detecting anomalies and unusual patterns in the time series data, which helps identify significant deviations from the norm.

```{r}

# Calculate the rolling mean and standard deviation with a window size of 12 months
library(zoo)
rolling_mean <- rollmean(AirPassengers, k = 12, fill = NA)
rolling_sd <- rollapply(AirPassengers, width = 12, FUN = sd, fill = NA)

# Identify months with passengers greater than mean + 2*SD (anomalies)
anomalies <- AirPassengers > (rolling_mean + 2 * rolling_sd)

# Plot the original data
plot(AirPassengers, 
     main = "Anomaly Detection in AirPassengers Data", 
     ylab = "Number of Passengers", 
     xlab = "Time", 
     col = "blue", 
     lwd = 2)

# Highlight the anomalies in red
points(time(AirPassengers)[anomalies], AirPassengers[anomalies], col = "red", pch = 19, cex = 1.5)

# Add a legend to distinguish between regular data points and anomalies
legend("topleft", 
       legend = c("Original Data", "Anomalies"), 
       col = c("blue", "red"), 
       pch = c(NA, 19), 
       lty = c(1, NA),
       lwd = 2)
```

<br> 
<br>

# Further Resources

Learn more about [package, technique, dataset] with the following:

<br>

* Resource I [Time Series Analysis in R](https://otexts.com/fpp3/)

* Resource II [R Documentation for AirPassengers Dataset](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/AirPassengers.html)

* Resource III [HoltWinters Exponential Smoothing in R](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/HoltWinters)

<br>
<br>

# Works Cited

This code through references and cites the following sources:

<br>

* Hyndman, R.J., & Athanasopoulos, G. (2018). Forecasting: Principles and Practice. [OTexts](https://otexts.com/fpp3/)

* R Core Team (2021). R: A Language and Environment for Statistical Computing. [R Foundation for Statistical Computing](https://www.R-project.org/)

* Cleveland, W.S., & Tiao, G.C. (1976). Decomposition of Seasonal Time Series: A Model for the Census X-11 Program. [Journal of the American Statistical Association](https://www.jstor.org/stable/2286010)

<br>
<br>